---
title: "Simulating acyclic pedigrees"
author: "Eric C. Anderson"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simulating acyclic pedigrees}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

So, my original approach to doing this was a fail because it actually produced
a whole lot of intergenerational marriage chains!  It is not enough to
just make sure that two mates don't share ancestors---they both must also
belong to different connected components.

I have a function in this package called `simulate_acyclic_pedigree()` that tries to
do that there.  I am still testing it.  

Here is what it looks like:
```{r}
library(tidyverse)
library(gdropR)
set.seed(555)
test <- simulate_acyclic_pedigree(Nm = 10, Nf = 10, end_time = 2)
```


That takes a long while (like, over a minute, with Nf = Nm = 40, and end_time = 5).
It is no doubt all the iteration being done to find full connected
components.

We will save that pedigree as FourGen_555_pedigree.rda.

If you want to visualize it you can do so like this (not evaluated by default):
```{r, eval=FALSE}
library(pedvis)

all_names <- unique(unlist(test$pedigree))

out3 <- ped2dot(
test$pedigree,
pfactorNodeStyle = "invis",
pfactorEdgeStyle = "invis",
ShowLabelNodes = all_names,
outf = "ped2dot_ex3"
)


```


That is not very pretty.

We should be able to test if it is acyclic by testing whether it is
a tree.  Theorem: an undirected graph is a tree if it is connected and
has N vertices and N-1 edges.



So, let's make a node list and an edgelist that we can read into iGraph.
```{r}
library(igraph)
# first translate it into a from->to data frame to read into
# igraph.  We will make it a factor graph, or course
tmp <- test$pedigree %>%
  mutate(factor_node = paste0(Pa, "-%-", Ma))

pa_down <- tmp %>% 
  group_by(Pa, factor_node) %>%
  slice(1) %>%
  select(Pa, factor_node) %>%
  rename(from = Pa, to = factor_node) %>%
  ungroup()

ma_down <- tmp %>% 
  group_by(Ma, factor_node) %>%
  slice(1) %>%
  select(Ma, factor_node) %>%
  rename(from = Ma, to = factor_node) %>%
  ungroup()
  
node_down <- tmp %>% 
  group_by(factor_node, Kid) %>%
  slice(1) %>%
  select(factor_node, Kid) %>%
  rename(from = factor_node, to = Kid) %>%
  ungroup()

graph_df <- bind_rows(
  pa_down,
  ma_down,
  node_down
)

g <- graph_from_data_frame(
  graph_df, 
  directed = FALSE
)

CC <- components(g)

# get a list of the vertices in each component
vert_list <- split(names(CC$membership), CC$membership)

# pull each of those components out as a separate graph
v_and_e_counts <- lapply(
  vert_list, 
  function(v) {
    g2 <- induced_subgraph(g, v)
    tibble(
      num_vertices = vcount(g2),
      num_edges = gsize(g2)
    )
  }
) %>%
  bind_rows() %>%
  mutate(diff = num_vertices - num_edges)

if(all(v_and_e_counts$diff == 1)) {
  print("Yep! It's a Tree!")
}
```
